---
title: "Identify flow direction with shp2graph"
author: "Jonathan Jupke"
date: "20 10 2020"
output: html_document
---

```{r setup, include=FALSE}

pacman::p_load(igraph,
               data.table,
               dplyr,
               here,
               kableExtra,
               magrittr,
               shp2graph,
               sf,
               sp,
               tmap)

knitr::opts_chunk$set(echo = TRUE, cache=FALSE)
pd.dir = "../../003_processed_data/"
rd.dir = "../../001_raw_data/"
tmap_mode("view")
```
# Load Data 
```{r load data}
dt_rivers = readRDS(
        file.path(pd.dir, "map_passability", "fixed_rivers.RDS")
)
st_sampling_sites = readRDS(
        file.path(rd.dir, "probestellen", "2020-09-23_all_sites.RDS")
)
```
# Map 
```{r map, fig.height=3, fig.width=4}
tm_shape(st_sampling_sites) + tm_dots()
```
I will use HR_06 and HR_05 for a test run.
```{r}
num_site_id1 = which(st_sampling_sites$site=="HR_05")
num_site_id2 = which(st_sampling_sites$site=="HR_06")
```
# Carpet data 
Here if find the stream segment that are closest to the chosen sampling points. **int_start** contains the ecoserv_id of the start segement and **in_end** that of the end segement. 
```{r carpet}
st_rivers=st_as_sf(dt_rivers)
st_sampling_sites %<>% st_transform (crs = st_crs(st_rivers))
int_start = st_nearest_feature(x = st_sampling_sites[num_site_id1, ],
                                       y = st_rivers)
int_start = st_rivers[int_start, ]$ecoserv_id
int_end = st_nearest_feature(x = st_sampling_sites[num_site_id2, ],
                                       y = st_rivers)
int_end = st_rivers[int_end,]$ecoserv_id
```

# First step to network 
I extract the length of the stream segements as weights for the network in **edge_length**. For further commands from the *shp2graph* package the spatial data needs to be in a spatial data frame (i.e. *sp*) format.  The derivation takes some time so I skip this here and load the saved file and hence the transformation is also commented out.  
```{r first network}
edge_length = st_length(st_rivers) %>% 
        as.numeric()

# st_rivers %<>% as_Spatial()

li_pre_net=readRDS(file.path(pd.dir, "network","2020-10-22_list_pre_network.RDS"))
igraph_river=readRDS(file.path(pd.dir, "network", "2020-10-22_igraph_river.RDS"))

```

Extract relevant information from the pre graph list. The new data table tb_edge_id holds the ids of segements, as well as the graph nodes and point_ids that belong to that stream/line/edge. 

```{r edge table}
tb_pre_join = tibble(EdgeID = li_pre_net[[3]][, 1],
                     Node1 = li_pre_net[[3]][, 2],
                     Node2 = li_pre_net[[3]][, 3])

tb_edge_id = left_join(x = tb_pre_join,
                       y = li_pre_net[[5]],
                       by = "EdgeID")

setDT(tb_edge_id)
```

Next we use the network in **igraph_river** to find the shortest route between the two points. 
```{r dijkstra}
# Node2 values of start and end points
from = tb_edge_id[ecoserv_id ==  int_start, Node2]
to = tb_edge_id[ecoserv_id ==  int_end, Node1]

# compute shortest path 
li_sp = shortest_paths(graph = igraph_river,
                       from = from,
                       to = to)
# extract nodes on path as vector 
v_path=as.vector(li_sp$vpath[[1]])

# create a subset of tb_edge_id that contains only segments with nodes in the
# path + start and end segement. This tibble will be used in the subsequent
# assignment of upstream and downstream movemen.
tb_edge_id_sub = tb_edge_id[(Node2 %in% v_path & Node1 %in% v_path)|ecoserv_id==int_start|ecoserv_id==int_end]
```

```{r show tb-edge-id-sub, echo=FALSE}
tb_edge_id_sub %>% 
  kable %>% 
  kable_minimal
```

The loop below identifies the flow direction within each segment as well as the order in which the segemnts are passed. 

```{r flow-direction-loop}
tb_edge_id_sub[, flow_direction := factor(levels=c("up","down"))]
tb_edge_id_sub[, segement_number := numeric()]

tb_edge_id_sub[ecoserv_id==int_start, c("flow_direction", "segement_number") := .("down",1) ]
# 
for (i in seq_along(v_path)) {
# for (i in 1:9){
#   i = 10
      loop_var = tb_edge_id_sub[is.na(flow_direction) &
                               (Node1 == v_path[i] | Node2 == v_path[i])]
      
        if (nrow(loop_var) == 1){
                tb_edge_id_sub[is.na(flow_direction) &
                               (Node1 == v_path[i] | Node2 == v_path[i]), segement_number := i]
          
          if (loop_var$FROM == tb_edge_id_sub[segement_number == (i -1), TO]) {
            tb_edge_id_sub[ecoserv_id == loop_var$ecoserv_id, flow_direction := "down"]
          } else if (loop_var$TO == tb_edge_id_sub[segement_number == (i - 1), TO]) {
            tb_edge_id_sub[ecoserv_id == loop_var$ecoserv_id, flow_direction := "up"]
          } else if (loop_var$TO == tb_edge_id_sub[segement_number == (i - 1), FROM]) {
            tb_edge_id_sub[ecoserv_id == loop_var$ecoserv_id, flow_direction := "up"]
          }
        }
        # any segments with this node at all?
        if (tb_edge_id_sub[is.na(flow_direction) &
                           (Node1 == v_path[i] | Node2 == v_path[i]), .N] == 0) {
                print(paste("No segments with Node", i, "left"))
                next()
        }

}

tb_edge_id_sub %>% 
  kable(format="html") %>% 
  kable_minimal
```

Now we can take the data table **dt_rivers** from the beginning and extract the relevant upstream and downstream passability scores for our river segements. 
```{r}
p_down = prod(dt_rivers[ecoserv_id %in% tb_edge_id_sub[flow_direction=="down", ecoserv_id], score_down])
p_up = prod(dt_rivers[ecoserv_id %in% tb_edge_id_sub[flow_direction=="up", ecoserv_id], score_down])
p_total = p_up*p_down
```
The code shown and discussed here is implemented for all sampling points in *


# System Information 
```{r session info, echo=FALSE}
Sys.time()
sessionInfo()
```

